Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe

### 2017-07-25 manuscript/usses_water_ms.Rmd
updates figs and a little writing

---
title: Linear ecosystem functional response across precipitation extremes in a sagebrush steppe
preprint: false
author: 
  - name: Andrew T. Tredennick
    affiliation: 1
    corresponding: true
    email: atredenn@gmail.com
  - name: Andrew R. Kleinhesselink
    affiliation: 2
  - name: Bret Taylor
    affiliation: 3
  - name: Peter B. Adler
    affiliation: 1
affiliation:
  - code: 1
    address: Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322
  - code: 2
    address: Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles, California 90095
  - code: 3
    address: USSES, address...
abstract: >
  Precipitation is predicted to become more variable in the western U.S., meaning years of above and below average precipitation will become more common. Such periods of drought and deluge are major drivers of plant community dynamics and ecosystem functioning in water limited grasslands. Here we report the results of a five-year experiment where we used drought and irrigation treatments (50% decrease/increase) to see how a sagebrush steppe plant community in Idaho will respond to future climate changes. Ecosystem functional response, defined as the relationship between soil moisture and aboveground net primary productivity, was suprisingly consistent across drought, control, and irrigation treatments. Aboveground net primary productivity (ANPP) responded positively to availabile soil moisture, but the response did not vary across treatments. There was also no evidence that treatment effects grew over time. The similarity of ecosystem functional response across treatment was not due to compensatory shifts at the plant community level, where species composition among treatments was similar and remarkably stable over the five years. At least in the short-term, ecosystem functional response and community composition in this sagebrush steppe system is resistant to increases and decreases in growing season precipitation.
header-includes: >
  \usepackage{lipsum}
  \usepackage{setspace}
  \usepackage{todonotes}
  \usepackage{rotating}
  \usepackage{color, soul}
  \usepackage{sectsty}
  \usepackage{bm,mathrsfs}
bibliography: ~/Dropbox/Bibliography/usses_water.bib
output: 
  rticles::peerj_article:
    base_format: rmarkdown::pdf_document # bookdown::pdf_document2 # for using \@ref()
---

@@

@@

Cumulative growing season soil moisture had a positive effect on ANPP (mean of $\beta_{1}$ = 0.43; 80\% BCI = -0.05, 0.88; 95\% BCI = -0.36, 1.15) (Fig. 2D).
Ecosystem functional response was similar among treatments, with treatment level regressions between soil moisture and log(ANPP) having similar intercepts (Fig. 3A) and slopes (Fig. 3B).
There was also no evidence that the treatment effects became more important over time because there was no directional trend in the random year effects (Fig. 4).

@@

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[height=5in]{../figures/Figure1.png}
  \caption{(A) Probability density of historical precipitation from 1926-2016, with the years of the experiment shown with arrows on the \emph{x}-axis. (B) Observed soil volumetric water content (VWC) over the course of the experiment. (C) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment. (D) Scatterplot of ANPP versus growing season precipitation. Colored regression lines are independently-fit linear models for each treatment with no random effects structure; dark black regression line is a linear regression through all the point with no random effects structure. Our analysis seeks to find if the data supports a model with different intercepts and/or slopes for each treatment, while also accounting for nonindependence of samples within plots across years. Color mapping for panels B-D is shown in the legend for panel C.}
\end{figure}

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[width=5in]{../figures/glmm_treatment_posteriors.png}
  \caption{Posterior distributions of treatment-level parameters ('Intercept' and the effect of 'Soil Moisture'). Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity.}
\end{figure}

>>>>

---
title: Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe
preprint: false
author: 
  - name: Andrew T. Tredennick
    affiliation: 1
    corresponding: true
    email: atredenn@gmail.com
  - name: Andrew R. Kleinhesselink
    affiliation: 2
  - name: Bret Taylor
    affiliation: 3
  - name: Peter B. Adler
    affiliation: 1
affiliation:
  - code: 1
    address: Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322
  - code: 2
    address: Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles, California 90095
  - code: 3
    address: USSES, address...
abstract: >
  Precipitation is predicted to become more variable in the western U.S., meaning years of above and below average precipitation will become more common. Such periods of drought and deluge are major drivers of plant community dynamics and ecosystem functioning in water limited grasslands. Here we report the results of a five-year experiment where we used drought and irrigation treatments (50% decrease/increase) to see how a sagebrush steppe plant community in Idaho will respond to future climate changes. Ecosystem functional response, defined as the relationship between soil moisture and aboveground net primary productivity, was suprisingly consistent across drought, control, and irrigation treatments. Aboveground net primary productivity (ANPP) responded positively to availabile soil moisture, but the response did not vary across treatments. There was also no evidence that treatment effects grew over time. The similarity of ecosystem functional response across treatment was not due to compensatory shifts at the plant community level, where species composition among treatments was similar and remarkably stable over the five years. At least in the short-term, ecosystem functional response and community composition in this sagebrush steppe system is resistant to increases and decreases in growing season precipitation.
header-includes: >
  \usepackage{lipsum}
  \usepackage{setspace}
  \usepackage{todonotes}
  \usepackage{rotating}
  \usepackage{color, soul}
  \usepackage{sectsty}
  \usepackage{bm,mathrsfs}
bibliography: ~/Dropbox/Bibliography/usses_water.bib
output: 
  rticles::peerj_article:
    base_format: rmarkdown::pdf_document # bookdown::pdf_document2 # for using \@ref()
---

@@

```{r get_results, echo=FALSE}
betas <- read.csv("../results/beta_summaries.csv")
```

@@

Cumulative growing season soil moisture had a positive effect on ANPP (mean of $\beta_{1}$ = `r round(betas$mean[8],2)`; 80\% BCI = `r round(betas$X10.[8],2)`, `r round(betas$X90.[8],2)`; 95\% BCI = `r round(betas$X2.5.[8],2)`, `r round(betas$X97.5.[8],2)`) (Fig. 3).
Ecosystem functional response was similar among treatments, with treatment level regressions between soil moisture and log(ANPP) having similar intercepts and slopes (Fig. 3A).
There was also no evidence that the treatment effects became more important over time because there was no directional trend in the random year effects (Fig. 4).

@@

# AUTHOR CONTRIBUTIONS
\begin{itemize}
  \item Andrew T. Tredennick collected data, analyzed the data, wrote the paper, prepared figures and/or tables, reviewed
drafts of the paper.
  \item Andrew R. Kleinhesselink conceived and designed the experiments, performed the experiments, collected data, reviewed drafts of the paper.
  \item Peter B. Adler conceived and designed the experiments, performed the experiments, collected data, analyzed the data, reviewed
drafts of the paper.
\end{itemize}

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[height=5in]{../figures/Figure1.png}
  \caption{(A) Probability density of historical precipitation from 1926-2016, with the years of the experiment shown with arrows on the \emph{x}-axis. (B) Observed soil volumetric water content (VWC) over the course of the experiment. (C) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment.}
\end{figure}

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[width=5in]{../figures/glmm_main_results.png}
  \caption{Results from the generalized linear mixed effects model. (A) Posterior distributions of treatment-level parameters ('Intercept' and the effect of 'Soil Moisture'). Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity. (B) Scatterplot of the data and model estimates shown a solid lines. Model estimates come from treatment level coefficients (colored lines) and the overall mean coefficients (heavy black line). Note that we show log(ANPP) on the \emph{y}-axis of panel B; this same plot can be seen on the arithmetic scale in supporting material Fig. S2-1.}
\end{figure}

### 2017-07-25 manuscript/usses_water_ms.tex
updates figs and a little writing

\title{Consistent ecosystem functional response during five years of drought
and deluge in a sagebrush steppe}

@@

@@

\affil[1]{Department of Wildland Resources and the Ecology Center, Utah State
University, Logan, Utah 84322}
\affil[2]{Department of Ecology and Evolutionary Biology, University of
California, Los Angeles, Los Angeles, California 90095}

@@

% 
\begin{abstract}
Precipitation is predicted to become more variable in the western U.S.,
meaning years of above and below average precipitation will become more
common. Such periods of drought and deluge could become major drivers of
plant community dynamics and ecosystem functioning in water limited
grasslands. Here we report the results of a five-year experiment where
we used drought and irrigation treatments (50\% decrease/increase) to
see how a sagebrush steppe plant community in Idaho will respond to
future climate changes. Ecosystem functional response, defined as the
relationship between soil moisture and aboveground net primary
productivity, was suprisingly consistent across drought, control, and
irrigation treatments. Aboveground net primary productivity (ANPP)
responded positively to availabile soil moisture, but the response did
not vary across treatments. There was also no evidence that treatment
effects grew over time. The similarity of ecosystem functional response
across treatment was not due to compensatory shifts at the plant
community level, where species composition among treatments was similar
and remarkably stable over the five years. At least in the short-term,
ecosystem functional response and community composition in this
sagebrush steppe system is resistant to increases and decreases in
growing season precipitation.
% Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text.
\end{abstract}

@@

Cumulative growing season soil moisture had a positive effect on ANPP
(mean of \(\beta_{1}\) = 0.43; 80\% BCI = -0.05, 0.88; 95\% BCI = -0.36,
1.15) (Fig. 2D). Ecosystem functional response was similar among
treatments, with treatment level regressions between soil moisture and
log(ANPP) having similar intercepts (Fig. 3A) and slopes (Fig. 3B).
There was also no evidence that the treatment effects became more
important over time because there was no directional trend in the random
year effects (Fig. 4).

@@

>>>>

\title{Consistent ecosystem functional response across precipitation extremes
in a sagebrush steppe}

@@

\author[3]{Bret Taylor}

@@

\affil[1]{Department of Wildland Resources and the Ecology Center, Utah State
University, Logan, Utah 84322}
\affil[2]{Department of Ecology and Evolutionary Biology, University of
California, Los Angeles, Los Angeles, California 90095}
\affil[3]{USSES, address\ldots{}}

@@

% 
\begin{abstract}
Precipitation is predicted to become more variable in the western U.S.,
meaning years of above and below average precipitation will become more
common. Such periods of drought and deluge are major drivers of plant
community dynamics and ecosystem functioning in water limited
grasslands. Here we report the results of a five-year experiment where
we used drought and irrigation treatments (50\% decrease/increase) to
see how a sagebrush steppe plant community in Idaho will respond to
future climate changes. Ecosystem functional response, defined as the
relationship between soil moisture and aboveground net primary
productivity, was suprisingly consistent across drought, control, and
irrigation treatments. Aboveground net primary productivity (ANPP)
responded positively to availabile soil moisture, but the response did
not vary across treatments. There was also no evidence that treatment
effects grew over time. The similarity of ecosystem functional response
across treatment was not due to compensatory shifts at the plant
community level, where species composition among treatments was similar
and remarkably stable over the five years. At least in the short-term,
ecosystem functional response and community composition in this
sagebrush steppe system is resistant to increases and decreases in
growing season precipitation.
% Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text.
\end{abstract}

@@

Cumulative growing season soil moisture had a positive effect on ANPP
(mean of \(\beta_{1}\) = 0.41; 80\% BCI = -0.08, 0.86; 95\% BCI = -0.45,
1.13) (Fig. 3). Ecosystem functional response was similar among
treatments, with treatment level regressions between soil moisture and
log(ANPP) having similar intercepts and slopes (Fig. 3A). There was also
no evidence that the treatment effects became more important over time
because there was no directional trend in the random year effects (Fig.
4).

@@

\section{AUTHOR CONTRIBUTIONS}\label{author-contributions}

\begin{itemize}
  \item Andrew T. Tredennick collected data, analyzed the data, wrote the paper, prepared figures and/or tables, reviewed
drafts of the paper.
  \item Andrew R. Kleinhesselink conceived and designed the experiments, performed the experiments, collected data, reviewed drafts of the paper.
  \item Peter B. Adler conceived and designed the experiments, performed the experiments, collected data, analyzed the data, reviewed
drafts of the paper.
\end{itemize}
